Manifold fitting by ridge estimation: a subspace-constrained approach

### 2021-02-25 Ridge_Estimation_main/Ridge-final.tex
zzh

\begin{assumption}\label{Data_assumption}
Assume there is a density function $p(x)$ such that 

@@

\end{assumption}

@@

Using assumption \ref{Data_assumption} and Lemma \ref{margin}, we have, for any $\hat{x}_k\in \hat{\mathcal G}$, a $\tilde{y}\in {\mathcal M}$ that is the end point of the unit speed curve $\gamma(t)$ starting from $\hat{x}_k$. Meanwhile, both  $\hat{x}_k$ and $\tilde{y}$ satisfy the ridge definition in \eqref{ridgeDF}. 
Thus, we know that 
\begin{equation}\label{pointwise-bound}
\min_{\tilde{y}_s\in {\mathcal M}_{\hat{\mathcal G}}}\|\hat{x}_k-\tilde{y}_s\|_2 \leq \|\hat{x}_k-\tilde{y}\|_2 = O (\|H_{\hat{p}}(\hat{x}_k)-H(\hat{x}_k)\|_F+\|g_{\hat{p}}(\hat{x}_k)-g(\hat{x}_k)\|_2  ).
\end{equation}
\begin{remark}
The distance between two discrete sets $\hat{\cal G}, {\cal M}_{\hat{\cal G}}$ is converted to the distance between $\hat{\cal G}$ and ${\cal M}$ by the unit speed curve $\gamma(t)$. $\gamma(t)$ starts from a point in $\hat{\cal G}$ (such as $\hat{x}_k$) with a direction of $\Pi_{H_p}(\gamma(t)) g_{p}(\gamma(t))$. The point at which $\gamma(t)$ goes through $\cal M$ is denoted by $\tilde{y}$. The details of the construction method of the curve $\gamma(t)$ can be found in the supplementary material.
%appendix \ref{Ridge Derivative Lemma}.
\end{remark}
From \eqref{Hausd}, by maximizing the distance with respect to $x_k\in 
\hat{\cal G}$ in \eqref{pointwise-bound}, we know that ${\rm Haus}({\mathcal {\hat{G}}}, {\mathcal M}_{\mathcal {\hat{G}}})$ yields a uniform upper bound by the sup norm at the right side:
\begin{equation}\label{Hausmargin}
{\rm Haus}({\hat{\mathcal G}}, {\mathcal M}_{\hat{\mathcal G}}) = O (\sup_{\hat{x}\in\hat{\cal G}}( \|H_{\hat{p}}(\hat{x})-H(\hat{x})\|_F+\|g_{\hat{p}}(\hat{x})-g(\hat{x})\|_2 ) ),
\end{equation}
where $\hat{p}$ is the estimated density, which can be chosen as $\hat{p}_h(x)$, $\hat{p}_{r,h}(x)$, or any other empirical forms.

@@

\begin{assumption}\label{assum}
There is a hyper-parameter $r_1$ such that $\mu(r_1,h)$ yields a uniform upper bound $\ell\in(0,1)$.
\end{assumption}
\remark{
Using this assumption, we can obtain a result on the margin or Hausdorff distance with respect to the estimated ridge. When $\mu(r_1,h)$ has a uniform bound $\ell$ for any $h$, we can neglect the specific form of $\mu(r_1,h)$ and obtain a result for the worst occasions (i.e. the most conservative result).

@@

We have two functions $\nu(h) = a_0 h^2 + a_1 \sqrt{\frac{1}{nh^{D+m}}}$ and $\nu_\ell (h) = \ell a_0 h^2 + a_1 \sqrt{\frac{1}{nh^{D+m}}}$ with $m=2,4,\ell \in(0,1)$. Then, the optimal minimums of them have the following relationship:  $\min_h \nu_\ell (h) = \ell^{\frac{D+2}{D+6}}\min_h \nu(h)$

@@

\begin{theorem} If the Assumption \ref{assum} is satisfied and the optimal $h$ is chosen with respect to the sample size $n$, %there is $r_1$ such that $\mu(r_1, h)\leq \ell,\ell\in(0,1)$, 

@@

{\rm Haus}(\hat{\cal G}_{{\it l-}{\rm SCRE}}, {\cal M}_{\hat{\cal G}_{{\it l}-{\rm SCRE}}}) \leq \ell^{\frac{D+2}{D+6}} {\rm Haus}(\hat{\cal G}_{\rm SCRE}, {\cal M}_{\hat{\cal G}_{\rm SCRE}}).

@@

@@

It is clear from Table \ref{result_compare} and Figure \ref{fig:circle6} that, under the same parameter setting (such as $h$ and noise level $\sigma$), the algorithm of {\it l}-SCRE yields a ridge with a margin much smaller than that of SCRE. The rate of increase of the distance measurement for {\it l}-SCRE is much lower than that for SCRE with the increasing $h$. This phenomenon is also explained in our theoretical analysis.

@@

\begin{table}[h]
\caption{The margin and Hausdorff between $\cal \hat{G}$ and $\mathcal M$ vary with $h$ for SCRE, {\it l}-SCRE, MFIT-i and MFIT-ii on the 2-dimensional sphere}.
\resizebox{14cm}{!}{
\label{result_compare}
\begin{tabular}{c|c|ccccccccc}\hline
  & $h$   & 0.1   &0.2 &0.3 &0.4&0.5&0.6&0.7&0.8&0.9\\\hline
  \multirow{4}{*}{Marg}  &SCRE 
& 0.0296 & 0.0304 & 0.0557 & 0.0955 & 0.1552 & 0.2478 & 0.4155 & 0.7002 & 0.3371 \\
  &{\it l}-SCRE
& 0.0274 & 0.0272 & 0.0323 & 0.0348 & 0.0360 & 0.0366 & 0.0371 & 0.0373 & 0.0375 \\
  &MFIT-i
& 0.1117 & 0.0409 & 0.0305 & 0.0228 & 0.0305 & 0.0457 & 0.0639 & 0.0849 & 0.1098 \\
  &MFIT-ii
& 0.0288 & 0.0357 & 0.0332 & 0.0285 & 0.0327 & 0.0444 & 0.0591 & 0.0763 & 0.0951 \\  \hline
\multirow{4}{*}{Haus}  &SCRE
& 0.1485 & 0.0891 & 0.0977 & 0.1303 & 0.1915 & 0.3047 & 0.5106 & 0.8972 & 0.9847 \\
  &{\it l}-SCRE
& 0.2012 & 0.0915 & 0.0799 & 0.0790 & 0.0800 & 0.0817 & 0.0827 & 0.0833 & 0.0837 \\
  &MFIT-i
& 0.9114 & 0.4856 & 0.1323 & 0.1129 & 0.0761 & 0.0849 & 0.0990 & 0.1191 & 0.1443 \\
  &MFIT-ii
& 0.1117 & 0.1201 & 0.1384 & 0.1648 & 0.1280 & 0.1146 & 0.1360 & 0.1417 & 0.1786 \\
 \hline
\end{tabular}}
\end{table}

In Table \ref{result_compare}, we provide another randomly generated circle-sample dataset. As shown in the table, when $h$ is too small (for instance in the cases shown by `-'), we cannot ensure that the domain of $x$ contains enough samples to run both MFIT-i and MFIT-ii. By comparing the margin and Hausdorff distance, we can conclude that the result of {\it l}-SCRE is more stable and competitive than the results of related methods.

>>>>

\begin{claim}\label{Data_assumption}
For any lower-dimensional manifold $\cal M$, there is a corresponding density function $p(x)$ such that 

@@

\end{claim}

@@

Using Claim \ref{Data_assumption} and Lemma \ref{margin}, we have, for any $\hat{x}_k\in \hat{\mathcal G}$, a $\tilde{y}\in {\mathcal M}$ that is the end point of the unit speed curve $\gamma(t)$ starting from $\hat{x}_k$. Meanwhile, both  $\hat{x}_k$ and $\tilde{y}$ satisfy the ridge definition in \eqref{ridgeDF}. 
Thus, we know that 
\begin{equation}\label{pointwise-bound}
\min_{\tilde{y}_s\in {\mathcal M}_{\hat{\mathcal G}}}\|\hat{x}_k-\tilde{y}_s\|_2 \leq \|\hat{x}_k-\tilde{y}\|_2 = O (\|H_{\hat{p}}(\hat{x}_k)-H(\hat{x}_k)\|_F+\|g_{\hat{p}}(\hat{x}_k)-g(\hat{x}_k)\|_2  ).
\end{equation}
\begin{remark}
The distance between two discrete sets $\hat{\cal G}, {\cal M}_{\hat{\cal G}}$ is converted to the distance between $\hat{\cal G}$ and ${\cal M}$ by the unit speed curve $\gamma(t)$. $\gamma(t)$ starts from a point in $\hat{\cal G}$ (such as $\hat{x}_k$) with a direction of $\Pi_{H_p}(\gamma(t)) g_{p}(\gamma(t))$. The point at which $\gamma(t)$ goes through $\cal M$ is denoted by $\tilde{y}$. The details of the construction method of the curve $\gamma(t)$ can be found in the supplementary material.
%appendix \ref{Ridge Derivative Lemma}.
\end{remark}
From \eqref{Hausd}, by maximizing the distance with respect to $x_k\in 
\hat{\cal G}$ in \eqref{pointwise-bound}, we know that ${\rm Haus}({\mathcal {\hat{G}}}, {\mathcal M}_{\mathcal {\hat{G}}})$ yields a uniform upper bound by the sup norm at the right side:
\begin{equation}\label{Hausmargin}
{\rm Haus}({\hat{\mathcal G}}, {\mathcal M}_{\hat{\mathcal G}}) = O (\sup_{\hat{x}\in\hat{\cal G}}( \|H_{\hat{p}}(\hat{x})-H(\hat{x})\|_F+\|g_{\hat{p}}(\hat{x})-g(\hat{x})\|_2 ) ),
\end{equation}
where $\hat{p}$ is the estimated density, which can be chosen as $\hat{p}_h(x)$, $\hat{p}_{r,h}(x)$, or any other empirical forms.

@@

%\begin{assumption}\label{assum}
%There is a hyper-parameter $r_1$ such that $\mu(r_1,h)$ yields a uniform upper bound $\ell\in(0,1)$.
%\end{assumption}
%\remark{
%Using this assumption, we can obtain a result on the margin or Hausdorff distance with respect to the estimated ridge. When $\mu(r_1,h)$ has a uniform bound $\ell$ for any $h$, we can neglect the specific form of $\mu(r_1,h)$ and obtain a result for the worst occasions (i.e. the most conservative result).
%}

@@

We have two functions $\nu(h) = a_0 h^2 + a_1 \sqrt{\frac{1}{nh^{D+m}}}$ and $\nu_\ell (h) = \ell a_0 h^2 + a_1 \sqrt{\frac{1}{nh^{D+m}}}$ with $m=2,4,\ell \in(0,1]$. Then, the optimal minimums of them have the following relationship:  $\min_h \nu_\ell (h) = \ell^{\frac{D+2}{D+6}}\min_h \nu(h)$.%, where $\ell = \sup_h \mu(r_1,h)\in (0,1]$.

@@

\begin{theorem}
 %If the Assumption \ref{assum} is satisfied and the optimal $h$ is chosen with respect to the sample size $n$, 
 If the optimal $h$ is chosen with respect to the sample size $n$, 
 %there is $r_1$ such that $\mu(r_1, h)\leq \ell,\ell\in(0,1)$, 

@@

{\rm Haus}(\hat{\cal G}_{{\it l-}{\rm SCRE}}, {\cal M}_{\hat{\cal G}_{{\it l}-{\rm SCRE}}}) \leq \ell^{\frac{D+2}{D+6}} {\rm Haus}(\hat{\cal G}_{\rm SCRE}, {\cal M}_{\hat{\cal G}_{\rm SCRE}}),

@@

where $\ell\in(0,1]$. Furthermore, if $r$ is selected as $r = r_1$ such that $\ell=\sup_h \mu(r_1, h)<1$, the factor will satisfy $\ell^{\frac{D+2}{D+6}}<1$ strictly.

@@

It is clear from Table \ref{result_compare}, Table \ref{result_compare2} and Figure \ref{fig:circle6} that, under the same parameter setting (such as $h$ and noise level $\sigma$), the algorithm of {\it l}-SCRE yields a ridge with a margin much smaller than that of SCRE. The rate of increase of the distance measurement for {\it l}-SCRE is much lower than that for SCRE with the increasing $h$. This phenomenon is also explained in our theoretical analysis.

@@

\begin{table}[h]
\caption{The margin and Hausdorff between $\cal \hat{G}$ and $\mathcal M$ vary with $h$ for SCRE, {\it l}-SCRE, MFIT-i and MFIT-ii on the 2-dimensional sphere}.
\resizebox{14cm}{!}{
\label{result_compare2}
\begin{tabular}{c|c|ccccccccc}\hline
  & $h$   & 0.1   &0.2 &0.3 &0.4&0.5&0.6&0.7&0.8&0.9\\\hline
  \multirow{4}{*}{Marg}  &SCRE 
& 0.0296 & 0.0304 & 0.0557 & 0.0955 & 0.1552 & 0.2478 & 0.4155 & 0.7002 & 0.3371 \\
  &{\it l}-SCRE
& 0.0274 & 0.0272 & 0.0323 & 0.0348 & 0.0360 & 0.0366 & 0.0371 & 0.0373 & 0.0375 \\
  &MFIT-i
& 0.1117 & 0.0409 & 0.0305 & 0.0228 & 0.0305 & 0.0457 & 0.0639 & 0.0849 & 0.1098 \\
  &MFIT-ii
& 0.0288 & 0.0357 & 0.0332 & 0.0285 & 0.0327 & 0.0444 & 0.0591 & 0.0763 & 0.0951 \\  \hline
\multirow{4}{*}{Haus}  &SCRE
& 0.1485 & 0.0891 & 0.0977 & 0.1303 & 0.1915 & 0.3047 & 0.5106 & 0.8972 & 0.9847 \\
  &{\it l}-SCRE
& 0.2012 & 0.0915 & 0.0799 & 0.0790 & 0.0800 & 0.0817 & 0.0827 & 0.0833 & 0.0837 \\
  &MFIT-i
& 0.9114 & 0.4856 & 0.1323 & 0.1129 & 0.0761 & 0.0849 & 0.0990 & 0.1191 & 0.1443 \\
  &MFIT-ii
& 0.1117 & 0.1201 & 0.1384 & 0.1648 & 0.1280 & 0.1146 & 0.1360 & 0.1417 & 0.1786 \\
 \hline
\end{tabular}}
\end{table}

%In Table \ref{result_compare}, we provide another randomly generated circle-sample dataset. 
As shown in Table \ref{result_compare} and Table \ref{result_compare2}, when the radius is too small, we cannot ensure that the domain which is centered at $x$ and with a radius of $h$, contains enough samples to run both MFIT-i and MFIT-ii. By comparing the margin and Hausdorff distance, we can conclude that the result of {\it l}-SCRE is more stable and competitive than the results of related methods.
